---
title: "Hands-on Exercise 10"
description: |
  A short description of the post.
author:
  - name: Darryl Kwok
    url: https://example.com/darrylkwok
date: 10-24-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r echo=TRUE, eval=TRUE}
packages = c('tmap', 'tidyverse',
             'sp', 'caret',
             'geojsonio',
             'reshape2', 'broom', 'sf', 'stplanr')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

The code below is used to install the appropriate version of sf

**Note that you only need to install once**

After installing, we will launch the library
```{r eval=FALSE}
library(devtools)
install_version("sf", version = "0.9-8", repos = "http://cran.us.r-project.org")

library(sf)
```


# Geospatial Data

In this section, you will download a copy of Greater Capital City Statistical Areas boundary layer from a dropbox depository by using *geojson_read()* of **geojsonio** package.

The code chunk used is shown below.

```{r}
Aus <- geojson_read("https://www.dropbox.com/s/0fg80nzcxcsybii/GCCSA_2016_AUST_New.geojson?raw=1", what = "sp")
```  

Lets extract the data by using the code chunk below.

```{r}
Ausdata <- Aus@data
```

The original data format is geojson, we will need to convert it to sf format and set the coordinates to be 4283

```{r}
AusSF <- st_as_sf(Aus) %>% 
  st_set_crs(4283) 
```

Check if the geometries are valid

```{r}
st_is_valid(AusSF)
```

The above output shows that there are several invalid features.

Let's fix them using the code chunk below.

```{r}
st_make_valid(AusSF)
```

```{r}
st_is_valid(AusSF)
```



### Displaying the boundary layer

Check if the boundary layer is correct.

```{r}
tmap_mode("plot")
qtm(AusSF) + 
  tmap_options(check.and.fix = TRUE)
```

### Displaying data table

You can view the simple feature data.frame by using the code chunk below.

```{r}
head(AusSF, 10)
```

Reorder the GCCSA_CODE order

```{r}
AusSF1 <- AusSF[order(AusSF$GCCSA_CODE),]
```

You can take a look at the data.frame again.

```{r}
head(AusSF1, 10)
```

### Converting into sp object

Convert AusSF1 from a dataframe into a sp object.

```{r}
Aus <- as(AusSF1, "Spatial")
```

## Calculating a distance matrix

In our spatial interaction model, space is one of the key predictor variables. In this example we will use a very simple Euclidean distance measure between the centroids of the Greater Capital City Statistical Areas as our measure of space.

**Caution note: With some areas so huge, there are obvious potential issues with this (for example we could use the average distance to larger settlements in the noncity areas), however as this is just an example, we will proceed with a simple solution for now.**

### Re-projecting to projected coordinate system

Perform coordinate projection to the correct projection system

```{r}
AusProj <- spTransform(Aus,"+init=epsg:3112")
summary(AusProj)
```

### Computing distance matrix

Use spDists to compute the distance matrix.

Although st_distance could be used, however in view of faster processing, we used spDists.

```{r}
dist <- spDists(AusProj)
dist 
```

### Converting distance matrix into distance pair list

In order to integrate the distance matrix with the migration flow data.frame later on, we need to transform the newly derived distance matrix into a three columns distance values list.

The code chunk below uses *melt()* of **reshape2** package of R to complete the task, however, you are encourage to archive the same task by using *pivot_longer()* of **dplyr** package. 

```{r}
distPair <- melt(dist)
head(distPair, 10)
```

### Converting unit of measurement from metres into km

The unit of measurement of Australia projected coordinate system is in meters.  As a result, the values in the distance matrix are in meters too.  The code chunk below is used to convert the distance values into kilometers.  

```{r}
distPair$value <- distPair$value / 1000
head(distPair, 10)
```

# Importing Interaction Data

Import the migration data into RStudio by using the code chunk below.

```{r}
mdata <- read_csv("https://www.dropbox.com/s/wi3zxlq5pff1yda/AusMig2011.csv?raw=1",col_names = TRUE)
glimpse(mdata)
```

## Combining the imported migration data

We need to add in our distance data that we generated earlier and create a new column of total flows which excludes flows that occur within areas (we could keep the within-area (intra-area) flows in, but they can cause problems so for now we will just exclude them). 

Firstly, create a new total column which excludes intra-zone flow totals. Then set them to a very small number to avoid causing the intra-zonal distance to become 0.

```{r}
mdata$FlowNoIntra <- ifelse(mdata$Orig_code == mdata$Dest_code,0,mdata$Flow)
mdata$offset <- ifelse(mdata$Orig_code == mdata$Dest_code,0.0000000001,1)
```

Since we have ordered the spatial data earlier so that the zones are in their code order. We can now easily join these data together.

```{r}
mdata$dist <- distPair$value 
```

Rather than setting the intra-zonal distances to 0, we should set them to something small (most intrazonal moves won't occur over 0 distance)

```{r}
mdata$dist <- ifelse(mdata$dist == 0,5,mdata$dist)
```

Let’s have a quick look at what your spangly new data looks like:

```{r}
glimpse(mdata)
```

# Visualising with desire line

In this section, you will learn how to prepare a desire line by using **stplanr** package.

## Removing intra-zonal flows

We will not plot the intra-zonal flows.  The code chunk below will be used to remove intra-zonal flows.

```{r}
mdatasub <- mdata[mdata$Orig_code!=mdata$Dest_code,]
```

First, use the *od2line()* function **stplanr** package to remove all but the origin, destination and flow columns.

```{r}
mdatasub_skinny <- mdatasub[,c(2,4,5)]
travel_network <- od2line(flow = mdatasub_skinny, zones = Aus)
```

Next, convert the flows to WGS84 projection.

```{r}
travel_networkwgs <- spTransform(travel_network,"+init=epsg:4326" )
```

Repeat the step for the Aus layer.

```{r}
AusWGS <- spTransform(Aus,"+init=epsg:4326" )
```

Lastly, we will set the line widths to some sensible value according to the flow.

```{r}
w <- mdatasub_skinny$Flow / max(mdatasub_skinny$Flow) * 10
```

Now, we are ready to plot the desire line map by using the code chunk below.

```{r}
plot(travel_networkwgs, lwd = w)
plot(AusWGS, add=T)
```


# Building Spatial Interaction Models

It is time for us to learn how to using R Stat function to calibrate the Spatial Interaction Models.  Instead of using *lm()* the [*glm()*](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/glm) function will be used. This is because *glm()* allow us to calibrate the model using generalised linear regression methods.

Note: Section 2.2.2 of Modelling population flows using spatial interaction models provides a detail discussion of generalised linear regression modelling framework.


## Unconstrained Spatial Interaction Model

In this section, we will calibrate an unconstrained spatial interaction model by using *glm()*.  The explanatory variables are origin population (i.e. vi1_origpop), destination median income (i.e. wj3_destmedinc) and distance between origin and destination in km (i.e. dist).

The code chunk used to calibrate to model is shown below:

```{r}
uncosim <- glm(Flow ~ log(vi1_origpop)+log(wj3_destmedinc)+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
summary(uncosim)
```

The model output report shows that the parameter estimates of the explanatory variables are significant at alpha value 0.001.

###  Fitting the model

To assess the performance of the model, we will use the *fitted()* of R to compute the fitted values.

```{r}
mdatasub$fitted <- fitted(uncosim)
```

### The more difficult ways (optional)

Another way to calculate the estimates is to plug all of the parameters back into Equation 6 like this:

First, assign the parameter values from the model to the appropriate variables

```{r}
k <- uncosim$coefficients[1]
mu <- uncosim$coefficients[2]
alpha <- uncosim$coefficients[3]
beta <- -uncosim$coefficients[4]
```

Next, plug everything back into the Equation 6 model... (be careful with the positive and negative signing of the parameters as the beta parameter may not have been saved as negative so will need to force negative)

```{r}
mdatasub$unconstrainedEst2 <- exp(k+(mu*log(mdatasub$vi1_origpop))+(alpha*log(mdatasub$wj3_destmedinc))-(beta*log(mdatasub$dist)))
```

which is exactly the same as this 

```{r}
mdatasub$unconstrainedEst2 <- (exp(k)*exp(mu*log(mdatasub$vi1_origpop))*exp(alpha*log(mdatasub$wj3_destmedinc))*exp(-beta*log(mdatasub$dist)))
```

### Saving the fitted values

Now, we will run the model and save all of the new flow estimates in a new column in the dataframe.

```{r}
mdatasub$unconstrainedEst2 <- round(mdatasub$unconstrainedEst2,0)
sum(mdatasub$unconstrainedEst2)
```

Next, we will turn the output into a little matrix by using [*dcast()*](https://www.rdocumentation.org/packages/maditr/versions/0.7.4/topics/dcast) of  **maditr** package.

```{r}
mdatasubmat2 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "unconstrainedEst2", margins=c("Orig_code", "Dest_code"))
mdatasubmat2
```

and compare with the original matrix by using the code chunk below.

```{r}
mdatasubmat <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "Flow", margins=c("Orig_code", "Dest_code"))
mdatasubmat
```

We can also visualise the actual flow and estimated flow by scatter plot technique.

```{r}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `unconstrainedEst2`))+
  geom_point(color="black", fill="light blue")
```


### Assessing the model performance

To provide a more formal assessment of the model, Goodness-o-Fit statistics will be used.  The code chunk below uses *postReSample()* of **caret** package to compute three Goodness-of-Fit statistics.

```{r}
postResample(mdatasub$Flow,mdatasub$unconstrainedEst2)
```

Notice that the R-squared value of 0.32 is relatively low.  It seems that the uncontrained model failed to fit the empirical data well.


## Origin Constrained Spatial Interaction Model

In this section, we will calibrate an origin constrained SIM (the "-1" indicates no intercept in the regression model) by using glm().

```{r}
origSim <- glm(Flow ~ Orig_code+log(wj3_destmedinc)+log(dist)-1, na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
#let's have a look at it's summary...
summary(origSim)
```

We can examine how the constraints hold for destinations this time.

Firstly, we will fitted the model and roundup the estimated values by using the code chunk below.

```{r}
mdatasub$origSimFitted <- round(fitted(origSim),0)
```

Next, we will used the step you had learned in previous section to create pivot table to turn paired list into matrix.

```{r}
mdatasubmat3 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "origSimFitted", margins=c("Orig_code", "Dest_code"))
mdatasubmat3
```

You can then compare with the original observed data as shown below.

```{r}
mdatasubmat
```

Next, let us display the actual flow and estimated flow by using the scatter plot technique.

```{r}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `origSimFitted`))+
  geom_point(color="black", fill="light blue")
```

Lastly, we compare the fitted values and the actual values by computing Goodness-of-fit statistics.

```{r}
postResample(mdatasub$Flow,mdatasub$origSimFitted)
```

Notice that the R-squared improved considerably from 0.32 in the unconstrained model to 0.43 in this origin constrained model.


## Destination Constrained Spatial Interaction Model

In this section, we will calibrate a destination constrained SIM (the "-1" indicates no intercept in the regression model) by using glm().

```{r}
destSim <- glm(Flow ~ Dest_code+log(vi1_origpop)+log(dist)-1, na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
summary(destSim)
```

We can examine how the constraints hold for destinations this time.  Firstly, we will fitted the model and roundup the estimated values by using the code chunk below.

```{r}
mdatasub$destSimFitted <- round(fitted(destSim),0)
```

Next, we will used the step you had learned in previous section to create pivot table to turn paired list into matrix.

```{r}
mdatasubmat6 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "destSimFitted", margins=c("Orig_code", "Dest_code"))
mdatasubmat6
```

Similar to the previous section, you can then compare with the original observed data as shown below.

```{r}
mdatasubmat
```

Next, let us display the actual flow and estimated flow by using the scatter plot technique.

```{r}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `destSimFitted`))+
  geom_point(color="black", fill="light blue")
```

Finally, we can test the Goodness-of-Fit in exactly the same way as before:

```{r}
postResample(mdatasub$Flow,mdatasub$destSimFitted)
```

Notice that the R-squared improved further from 0.32 in the unconstrained model to 0.65 in this origin constrained model.


## Doubly Constrained Spatial Interaction Model

In this section, we will calibrate a Doubly Constrained Spatial Interaction Model by using glm().


```{r}
doubSim <- glm(Flow ~ Orig_code+Dest_code+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
summary(doubSim)
```

We can examine how the constraints hold for destinations this time.  Firstly, we will fitted the model and roundup the estimated values by using the code chunk below.

```{r}
mdatasub$doubsimFitted <- round(fitted(doubSim),0)
```

Next, we will used the step you had learned in previous section to create pivot table to turn paired list into matrix.

```{r}
mdatasubmat7 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "doubsimFitted", margins=c("Orig_code", "Dest_code"))
mdatasubmat7
```

Similar to the previous section, you can then compare with the original observed data as shown below.

```{r}
mdatasubmat
```

Next, let us display the actual flow and estimated flow by using the scatter plot technique.

```{r}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `doubsimFitted`))+
  geom_point(color="black", fill="light blue")
```


