---
title: "Hands-on Exercise 6"
description: |
  A short description of the post.
author:
  - name: Darryl Kwok
    url: https://example.com/darrylkwok
date: 09-19-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE, eval=TRUE, echo=TRUE, message=FALSE, error=FALSE, fig.retina=3}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Data
```{r}
hunan <- st_read(dsn = "data/shapefile", 
                 layer = "Hunan")
```                 


```{r}
hunan2012 <- read_csv("data/attribute/Hunan_2012.csv")
```

## Join the Data

```{r}
hunan <- left_join(hunan,hunan2012)
```

```{r}
qtm(hunan, "GDPPC")
```

## Computing (QUEEN) contiguity based neighbours

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)

```

TO see the neighours for the first polygon in the object

```{r}
wm_q[[1]]
```

```{r}
hunan$County[1]
```

To reveal the county names of the five neighbouring polygons

```{r}
hunan$NAME_3[c(2,3,4,57,85)]
```

## Display the complete weight matrix by using str()

```{r}
str(wm_q)
```

## Creating (ROOK) contiguity based neighbours

```{r}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

## Visualising contiguity weights

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])

latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])

coords <- cbind(longitude, latitude)
```

### Check the first few observations to see if things are formatted correctly

```{r}
head(coords)
```

## Plotting Queen contiguity based neighbours maps

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```


## Plotting Rook contiguity based neighbours maps

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")

```

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red", main="Queen Contiguity")
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red", main="Rook Contiguity")

```



## Computing distance based neighbours

```{r}
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)

```

## Computing fixed distance weight matrix

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62

```

Use str() to display the content of wmd62 weight matrix

```{r}

str(wm_d62)

```


Another way to display the structure of the weigth matrix is to combine table() and card() of spdep

```{r}
table(hunan$County, card(wm_d62))
```

## Plotting fixed distance weight matrix

Plotting them on top of each other

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```

Plotting them side by side
```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(k1, coords, add=TRUE, col="red", length=0.08, main="1st nearest neighbours")
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6, main="Distance link")

```


## Computing adpative distance weight matrix

```{r}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6

```

Display content of the matrix
```{r}

str(knn6)

```

### Plotting distance based neighbours

```{r}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")

```


# Weights based on Inverse Distance Weight

Computing distances between areas by using nbdists()

```{r}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

## Row standardised weights matrix

```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

See the weight of the first polygon's four neighbors type

```{r}
rswm_q$weights[10]

```

Derive a row standardised distance weight matrix

```{r}
rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids

```

```{r}
rswm_ids$weights[1]

```

```{r}
summary(unlist(rswm_ids$weights))
```

# Application of Spatial Weight Matrix

## Spatial lag with row standardized weights

```{r}
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
GDPPC.lag
```

Append the spatially lag GDPPC values onto hunan SpatialPolygonDataFrame by using the code chunk below

```{r}
lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
lag.res <- as.data.frame(lag.list)
colnames(lag.res) <- c("NAME_3", "lag GDPPC")
hunan <- left_join(hunan,lag.res)
```

Display the joined table

```{r}
head(hunan)
```

Plot both the GDPPC and spatial lag GDPPC for comparison using the code chunk below

```{r}
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)
```

## Spatial Window Sum

Assign knn6 to a new variable, because we have to directly alter its structure to add the diagonal elements

```{r}
knn6a <- knn6

```

To add the diagonal element to the neighbour list, us include.self()

```{r}
include.self(knn6a)

```

Assign binary weight to the neighbour structure that includes the diagonal element

```{r}
binary.knn6 <- lapply(knn6a, function(x) 0*x+1)
binary.knn6[1]
```

Use nb2listw() and glist() to explicity assign weight values

```{r}
wm_knn6 <- nb2listw(knn6a, glist = binary.knn6, style = "B")

```

With the new weight structure, compute the lag variable with lag.listw()

```{r}
lag_knn6 <- lag.listw(wm_knn6, hunan$GDPPC)

```

Convert the lag variable listw object into a datatframe

```{r}
lag.list.knn6 <- list(hunan$NAME_3, lag.listw(wm_knn6, hunan$GDPPC))
lag_knn6.res <- as.data.frame(lag.list.knn6)
colnames(lag_knn6.res) <- c("NAME_3", "lag_sum GDPPC")

```

Join the lag_sum GDPPC map to hunan sf

```{r}
hunan <- left_join(hunan, lag_knn6.res)
```

Plot GDPPC and lag sum GDPPC next to each other

```{r}
gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)

```



















