---
title: "In-Class Exercise 4"
description: |
  A short description of the post.
author:
  - name: Darryl Kwok
    url: https://example.com/darrylkwok
date: 09-06-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Importing and Installing all required Packages
- sf is mainly for importing geospatial data
- raster is for handling raster-based geospatial data

```{r}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
  install.packages(p)
  }
library(p,character.only = T)
}
```

## Importing Geospatial Data

```{r}
sg_sf <- st_read(dsn = "data/shapefile", layer="CostalOutline")
mpsz_sf <- st_read(dsn = "data/shapefile", layer = "MP14_SUBZONE_WEB_PL")
```

## Importing Aspatial Data

This childcare.rds is retreived from the OneMap API.
However, the childcare data in hands-on exercise 4 is retrieved from data.gov.sg and the data format is in geojson format. Compare the differences and you will realise that geojson truncates the data and made the data hard to interpret.
```{r}
childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")
```

### Converting from aspatial to geospatial

Set crs to 3414 based on Singapore's National Projection System
```{r}
CHAS_sf <- st_as_sf(CHAS, coords=c("X_COORDINATE","Y_COORDINATE"),
                    crs=3414)
```

Since we are using Longtitude and Latitude, the correct crs is 4326.
Therefore after calling the st_as_sf() function, we perform a mathematical transformation by using st_transform() to transform the geometrical properties to 3414 projection
```{r}
childcare$Lat <- as.numeric(childcare$Lat)
childcare$Lng <- as.numeric(childcare$Lng)
childcare_sf <- st_as_sf(childcare, coords=c("Lng", "Lat"), crs=4326) %>% 
  st_transform(crs = 3414)
```

## Geospatial Data Wrangling

### Converting from sf to Spatial Classes
```{r}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

### Converting from Spatial classes to sp format
This will remove all the other columns, and only keep the geometric features and projection
```{r}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

### Converting from sp to spatstat ppp format
After you convert it to ppp format, you will lose the projections, so if you convert to raster, you need to define the projections
```{r}
childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")
```









