---
title: "Hands-on Exercise 5"
description: |
  A short description of the post.
author:
  - name: Darryl Kwok
    url: https://example.com/darrylkwok
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE, eval=TRUE, echo=TRUE, message=FALSE, error=FALSE, fig.retina=3}
knitr::opts_chunk$set(echo = FALSE)
```

## Installing and launching the R packages

```{r}
packages = c('sp', 'rgdal', 'spNetwork', 'tmap')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

## Importing the Data

```{r}
network <- readOGR(dsn="data/geospatial", 
                   layer="Punggol_St",
                   verbose = FALSE)
childcare <- readOGR(dsn="data/geospatial",
                     layer="Punggol_CC",
                     verbose = FALSE)
```

Examine the structure of the output SpatialDataFrame 

```{r eval=FALSE}
str(network)
str(childcare)
```

Assign EPSG code to the SpatialDataFrames

```{r}
childcare <-spTransform(childcare, CRS("+init=epsg:3414"))
network <- spTransform(network,CRS("+init=epsg:3414"))
```

Visualise the Data

```{r}
plot(network)
plot(childcare,add=T,col='red',pch = 19)
```

Visualise the Data interactively
```{r}
tmap_mode('view')
tm_shape(childcare)+
  tm_dots() +
tm_shape(network)+
  tm_lines()

```


# Netowrk Constrained KDE (NetKDE Analysis)

## Preparing the lixels objects

```{r}
lixels <- lixelize_lines(network,700,mindist = 350)
```


## Generating line centre points

```{r}
samples <- lines_center(lixels)
```

## Performing NetKDE
```{r}
densities <- nkde(network, 
                  events = childcare,
                  w = rep(1,nrow(childcare)),
                  samples = samples,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 5, #we aggregate events within a 5m radius (faster calculation)
                  sparse = TRUE,
                  verbose = FALSE)
```

## Visualising NetKDE

Before we can visualise the NetKDE values, we need to insert the computed density values

```{r}
samples$density <- densities
lixels$density <- densities
```

Since the SVY21 projection system is in metres, the scale of the computed density values are very small (i.e. 0.0000000005). Hence we rescale it from metres to kilometers
```{r}
samples$density <- samples$density*1000
lixels$density <- lixels$density*1000
```


```{r}
tmap_mode('view')
tm_shape(lixels)+
  tm_lines(col="density")+
tm_shape(childcare)+
  tm_dots()
```

# Network COnstrained G- and K- function Analysis




















