---
title: "Hands-on Exercise 3"
description: |
  A short description of the post.
author:
  - name: Darryl Kwok
    url: https://example.com/darrylkwok
date: 08-30-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Import neccessary Packages
```{r}
packages = c('sf', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Read  MP14_SUBZONE_WEB_PL Shape File
```{r echo=TRUE}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

### Display the contents of mpsz read in
``` {r echo=TRUE}
mpsz
```

## Read Singapore Residents by Planning Area/ Subzone, Age Group, Sex and Type of Dwelling csv file
``` {r echo=TRUE}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

### Display the contents of popdata
```{r echo=TRUE}
popdata
```

## Data Wrangling
### Create new columns such as YOUNG, ECONOMY ACTIVE, AGED, TOTAL AND DEPENDENCY
```{r echo=TRUE}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG,
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:11]) + rowSums(.[13:15])) %>%
  mutate(`AGED` = rowSums(.[16:21])) %>%
  mutate(`TOTAL` = rowSums(.[3:21])) %>%
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, `ECONOMY ACTIVE`, `AGED`, `TOTAL`, `DEPENDENCY`)

popdata2020
```


```{r}
popdata[1:3]


```


## Join 2 datasets together

### Convert values in popdata to uppercase
```{r echo=TRUE}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ),
            .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)

popdata2020
```

### Joining Attribute Data and Geospatial Data
```{r echo=TRUE}
mpsz_pop2020 <- left_join(mpsz, popdata2020, by = c("SUBZONE_N" = "SZ"))

mpsz_pop2020

```

## Write mpszpop2020 in r format
```{r echo=TRUE}
mpsz_pop2020_rds <- write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")

```


## Plotting a choropleth map with qtm()

```{r echo=TRUE}
tmap_mode("plot")
qtm(mpsz_pop2020, fill="DEPENDENCY")

```

##Plotting a choropleth map using tmap's elements
```{r echo=TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY", style = "quantile", palette = "Blues", title = "Dependency Ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size=1.2,
            legend.height=0.45,
            legend.width=0.35,
            frame=TRUE) +
  tm_borders(alpha=0.5) +
  tm_compass(type="8star",size=2)+
  tm_scale_bar() + 
  tm_grid(alpha=0.2)+
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))


```

## Drawing a Base Map
```{r echo=TRUE}
tm_shape(mpsz_pop2020) + 
  tm_polygons()


```

### Include parameters in tm_polygons()
``` {r echo=TRUE}
tm_shape(mpsz_pop2020) +
  tm_polygons("DEPENDENCY")
```

### Using tm_fill() instead of tm_polygons()
```{r echo=TRUE}
tm_shape(mpsz_pop2020) + 
  tm_fill("DEPENDENCY")

```

### Using tm_fill() and tm_borders()
```{r echo=TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY") +
  tm_borders(lwd=0.1, alpha = 1)

```

### Using tm_fill() and tm_borders() for data classification that utilises 5 classes
```{r echo=TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY", n=5, style="jenks") +
  tm_borders(alpha=0.5)
```

### Using tm_fill() and tm_borders() for data classification that utilises equal classification
```{r echo=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

## Getting the summary info for dependency
``` {r echo=TRUE}
summary(mpsz_pop2020$DEPENDENCY)
```

## Plotting Choropleth map with custom break after getting the summary info
``` {r echo=TRUE}
tm_shape(mpsz_pop2020) + 
  tm_fill("DEPENDENCY", breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

## Colour Scheme for tmap
``` {r echo=TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          n=6,
          style="quantile",
          palette= "Blues") +
  tm_borders(alpha=0.5)


```

### Reverse colour shading
```{r echo=TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style="quantile",
          palette="-Greens") +
  tm_borders(alpha=0.5)
``` 

### Change Map Style
```{r echo=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")

```

### Using Compass, scale bar and grid lines for design
```{r echo=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```


## Drawing Small Mutiple Choropleth Maps 
### Using tm_fill
``` {r echo=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

### Using tm_polygons
```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))

```

### Using tm_facets
```{r echo=TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

### Using tmap_arrange()
```{r echo=TRUE}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)


```


## Mapping Spatial Object that meets a criteria
```{r echo=TRUE}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```







