---
title: "In Class Exercise 3"
description: |
  A short description of the post.
author:
  - name: Darryl Kwok
    url: https://example.com/darrylkwok
date: 08-30-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

``` {r echo=FALSE}
packages <- c('sf', 'tidyverse', 'tmap')
for(p in packages) {
  if(!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}
```
## Read the mpszpop2020 data 
```{r echo=TRUE}
mpszpop2020 <- read_rds("data/rds/mpszpop2020.rds")
```

### Drop records with NA values
```{r echo=TRUE}
mpszpop2020a <- mpszpop2020 %>%
  drop_na()

```


### Visualising Extreme Values using Boxplots
```{r echo=TRUE}
ggplot(data=mpszpop2020, 
       aes(x = "", 
           y = AGED)) +
  geom_boxplot()
```

### Create get.var function
``` {r echo=TRUE}
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```



### Plotting a percentile map using tmap functions
``` {r echo=TRUE}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- get.var("DEPENDENCY", mpszpop2020a)
bperc <- quantile(var,percent)
tm_shape(mpszpop2020) +
  tm_polygons() +
tm_shape(mpszpop2020a) +
  tm_fill("DEPENDENCY",
          title="DEPENDENCY",
          breaks=bperc,
          palette="Blues",
          labels=c("< 1%", "1% - 10%",
                   "10% - 50%", 
                   "50% - 90%",
                   "90% - 99%", 
                   "> 99%"))  +
  tm_borders() +
  tm_layout(title = "Percentile Map", 
            title.position = c("right",
                               "bottom"))

```

### Create percentile mapping function
``` {r echo=TRUE}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam,df)
  bperc <- quantile(var,percent)
  tm_shape(mpszpop2020) +
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(title = mtitle, title.position = c("right","bottom"))
}

```


### Testing the percentile mapping function
``` {r echo=TRUE}
youngmap <- percentmap("YOUNG", mpszpop2020a)
agedmap <- percentmap("AGED", mpszpop2020a)

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### Creating the boxbreaks function
```{r echo=TRUE}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

### Test the boxbreak function
```{r echo=TRUE}
var <- get.var("AGED", mpszpop2020a) 
boxbreaks(var)
```


### Create the boxmap function
``` {r echo=TRUE}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(mpszpop2020) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(title = mtitle, 
            title.position = c("right",
                               "bottom"))
}
```


### Test the Boxmap function using AGED population
```{r echo=TRUE}
boxmap("ECONOMY ACTIVE", mpszpop2020a)

```

## Plotting Raw Rate Map

### Prepare the data
``` {r echo=TRUE}
mpszpop2020a <- mpszpop2020 %>%
  mutate(`AGED%` = (`AGED`
/`TOTAL`)*100) %>%
  filter(`AGED%` >= 0)
```

### Plot the map
``` {r echo=TRUE}
var <- get.var("AGED%", mpszpop2020a)
boxbreaks(var)
boxmap("AGED%",mpszpop2020a)
```




