---
title: "Hands-on Exercise 11"
description: |
  A short description of the post.
author:
  - name: Darryl Kwok
    url: https://example.com/darrylkwok
date: 11-01-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.retina = 3, warning = FALSE, message = FALSE)
```

# Introduction

In this hands-on exercise, you will gain hands-on experience on how to model geographical accessibility by using R's geospatial analysis packages.

## Learning Outcome

By the end of this hands-on exercise, you will be able:

- to import GIS polygon data into R and save them as simple feature data frame by using appropriate functions of sf package of R;
- to import aspatial data into R and save them as simple feature data frame by using appropriate functions of sf package of R;
- to computer accessibility measure by using Hansen's potential model and Spatial Accessibility Measure (SAM); and
- to visualise the accessibility measures by using tmap and ggplot2 packages.


## The data

Four data sets will be used in this hands-on exercise, they are:

- `MP14_SUBZONE_NO_SEA_PL`: URA Master Plan 2014 subzone boundary GIS data.  This data set is downloaded from data.gov.sg.
- `hexagons`: A 250m radius hexagons GIS data.  This data set was created by using [*st_make_grid()*](https://r-spatial.github.io/sf/reference/st_make_grid.html) of sf package.  It is in ESRI shapefile format.
- `ELDERCARE`: GIS data showing location of eldercare service.  [This data](https://data.gov.sg/dataset/eldercare-services) is downloaded from data.gov.sg.  There are two versions.  One in ESRI shapefile format.  The other one in Google kml file format.  For the purpose of this hands-on exercise, ESRI shapefile format is provided.
- `OD_Matrix`: a distance matrix in csv format.  There are six fields in the data file.  They are:

  + `origin_id`: the unique id values of the origin (i.e. `fid` of hexagon data set.),
  + `destination_id`: the unique id values of the destination (i.e. `fid` of `ELDERCARE` data set.),
  + `entry_cost`: the perpendicular distance between the origins and the nearest road),
  + `network_cost`: the actual network distance from the origin and destination, 
  + `exit_cost`: the perpendicular distance between the destination and the nearest road), and
  + `total_cost`: the summation of `entry_cost`, `network_cost` and `exit_cost`.

All the values of the cost related fields are in **metres**.

**Reminder:** *Except `MP14_SUBZONE_NO_SEA_PL` data set, the other three data set are specially prepared by Prof. Kam for teaching and research purpose.  Students taking IS415 Geospatial Analytics and Applications are allowed to use them for hands-on exercise purpose.  Please obtain formal approval from Prof. Kam if you want to use them for other courses or usage.*

#  Getting Started
## Installing and launching R packages


```{r echo=TRUE, eval=TRUE}
packages = c('tmap', 'SpatialAcc',
             'sf', 'ggstatsplot', 'reshape2',
             'tidyverse', 'Rmpfr')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```



#  Geospatial Data Wrangling  
## Importing geospatial data



```{r echo=TRUE, eval=TRUE}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 
```


## Updating CRS information

The code chunk below updates the newly imported *mpsz* with the correct ESPG code (i.e. 3414)

```{r, echo=TRUE, eval=TRUE}
mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)
```

After transforming the projection metadata, you can verify the projection of the newly transformed *mpsz_svy21* by using *st_crs()* of sf package.

The code chunk below will be used to varify the newly transformed *mpsz_svy21*. 

```{r echo=TRUE, eval=TRUE}
st_crs(mpsz)
```

Notice that the EPSG: is indicated as *3414* now.

## Cleaning and updating attribute fields of the geospatial data

```{r}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)
```

```{r}
hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```


# Apsaital Data Handling and Wrangling
## Importing Distance Matrix



```{r message=FALSE}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

## Tidying distance matrix


```{r}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```



```{r}
distmat_km<-as.matrix(distmat/1000)
```

# Modelling and Visualising Accessibility using Hansen Method
## Computing Hansen's accessibility


```{r}
acc_Hansen <- data.frame(ac(hexcen$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "Hansen"))
```

The default field name is very messy, we will rename it to `accHansen` by using the code chunk below.

```{r}
colnames(acc_Hansen) <- "accHansen"
```


Next, we will convert the data table into tibble format by using the code chunk below.

```{r}
acc_Hansen <- tbl_df(acc_Hansen)
```



```{r}
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```



```{r eval=FALSE}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "Hansen"))
colnames(acc_Hansen) <- "accHansen"
acc_Hansen <- tbl_df(acc_Hansen)
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

## Visualising Hansen's accessibility 

### Extracting map extend


```{r echo=TRUE, eval=TRUE}
mapex <- st_bbox(hexagons) #view extent
```


```{r fig.width=12, fig.height=8}
tmap_mode("plot")
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_fill(col = "accHansen",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: Hansen method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical graphic visualisation


```{r}
hexagon_Hansen <- st_join(hexagon_Hansen, mpsz, 
                          join = st_intersects)
```

Next, *ggplot()* will be used to plot the distribution by using boxplot graphical method.    

```{r}
ggplot(data=hexagon_Hansen, 
       aes(y = log(accHansen), 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```


# Modelling and Visualising Accessibility using KD2SFCA Method
## Computing KD2SFCA's accessibility

```{r}
acc_KD2SFCA <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "KD2SFCA"))
colnames(acc_KD2SFCA) <- "accKD2SFCA"
acc_KD2SFCA <- tbl_df(acc_KD2SFCA)
hexagon_KD2SFCA <- bind_cols(hexagons, acc_KD2SFCA)
```

## Visualising KD2SFCA's accessibility 

The code chunk below uses a collection of mapping fucntions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.  Notice that `mapex` is reused for *bbox* argument.

```{r fig.width=12, fig.height=8}
tmap_mode("plot")
tm_shape(hexagon_KD2SFCA,
         bbox = mapex) + 
  tm_fill(col = "accKD2SFCA",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: KD2SFCA method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical graphic visualisation

Now, we are going to compare the distribution of KD2CFA accessibility values by URA Planning Region.

Firstly, we need to add the planning region field into *hexagon_KD2SFCA* simple feature data frame by using the code chunk below.

```{r}
hexagon_KD2SFCA <- st_join(hexagon_KD2SFCA, mpsz, 
                          join = st_intersects)
```


```{r}
ggplot(data=hexagon_KD2SFCA, 
       aes(y = accKD2SFCA, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```


# Modelling and Visualising Accessibility using Spatial Accessibility Measure (SAM) Method
## Computing SAM accessibility


```{r}
acc_SAM <- data.frame(ac(hexagons$demand,
                         eldercare$capacity,
                         distmat_km, 
                         d0 = 50,
                         power = 2, 
                         family = "SAM"))
colnames(acc_SAM) <- "accSAM"
acc_SAM <- tbl_df(acc_SAM)
hexagon_SAM <- bind_cols(hexagons, acc_SAM)
```

## Visualising SAM's accessibility 

The code chunk below uses a collection of mapping fucntions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.  Notice that `mapex` is reused for *bbox* argument.

```{r fig.width=12, fig.height=8}
tmap_mode("plot")
tm_shape(hexagon_SAM,
         bbox = mapex) + 
  tm_fill(col = "accSAM",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: SAM method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 3),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical graphic visualisation


```{r}
hexagon_SAM <- st_join(hexagon_SAM, mpsz, 
                       join = st_intersects)
```

ggplot() will be used to plot the distribution by using boxplot graphical method.    

```{r}
ggplot(data=hexagon_SAM, 
       aes(y = accSAM, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```
