---
title: "Hands-on Exercise 4"
description: |
  A short description of the post.
author:
  - name: Darryl Kwok
    url: https://example.com/darrylkwok
date: 09-05-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

#Importing the spatial data
```{r echo=TRUE}
childcare_sf <- st_read("data/geospatial/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```
```{r echo=TRUE}
sg_sf <- st_read(dsn = "data/geospatial", layer="CostalOutline")
```

```{r echo=TRUE}
mpsz_sf <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

### Get the system information of these geospatial data
```{r echo=TRUE}
st_crs(childcare_sf)
```
Here we can see that both mpsz_sf and sg_sf are projected in svy21, but their EPSG is 9001 when it is supposed to be 3414
```{r echo=TRUE}
st_crs(mpsz_sf)
```
```{r echo=TRUE}
st_crs(sg_sf)
```
### Correcting the EPSG code
```{r echo=TRUE}
mpsz_sf <- st_set_crs(mpsz_sf, 3414)
sg_sf <- st_set_crs(sg_sf, 3414)
```

### Mapping Geospatial data sets using mpsz to plot Singapore's map and childcare_sf as the dots
```{r echo=TRUE}
tm_shape(mpsz_sf) +
  tm_polygons() +
tm_shape(childcare_sf) +
  tm_dots()
```
### Converting sf data frames to sp's spatial class
```{r echo=TRUE}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

### Converting the Spatial class into generic sp format
```{r echo=TRUE}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")

```

### Converting the generic sp format into spatstat's ppp format
```{r echo=TRUE}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```

### Plot childcare_ppp
```{r echo=TRUE}
plot(childcare_ppp)

```

Get a summary of childcare_ppp
```{r echo=TRUE}
summary(childcare_ppp)

```
### Handling Duplicated points

This confirms that there are duplicated points
```{r echo=TRUE}
any(duplicated(childcare_ppp))
```
Count the number of co-indicence points
```{r echo=TRUE}
multiplicity(childcare_ppp)

```

Find out how many locations have more than one point event
```{r echo=TRUE}
sum(multiplicity(childcare_ppp))

```

View the duplicate locations
```{r echo=TRUE}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)

tmap_mode('plot')

```
Use jittering to add a small perturbation to the duplicate points
```{r echo=TRUE}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)


```

Check the Duplicate points
```{r echo=TRUE}
any(duplicated(childcare_ppp_jit))

```

### Creating and plotting owin object
```{r echo=TRUE}
sg_owin <- as(sg_sp, "owin")

plot(sg_owin)
```
Get the summary of sg_owin
```{r echo=TRUE}
summary(sg_owin)
```

### Combining point events object and owin object
Extract childcare events that are located within Singapore
```{r echo=TRUE}
childcareSG_ppp = childcare_ppp[sg_owin]

summary(childcareSG_ppp)
```

Plot the childcare events 
```{r echo=TRUE}
plot(childcareSG_ppp)

```

### Computing Kernel Density Estimation using automatic bandwidth selection
```{r echo=TRUE}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 


plot(kde_childcareSG_bw)
```

Retrieve the bandwidth that is used to compute the kernel density estimation layer
```{r echo=TRUE}
bw <- bw.diggle(childcareSG_ppp)
bw

```
### Rescalling KDE value
Convert the unit of measurement from meter to kilometer
```{r echo=TRUE}
childcareSG_ppp.km <- rescale(childcareSG_ppp, 1000, "km")

kde_childcareSG.bw <- density(childcareSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG.bw)
```


### Working with different automatic bandwidth methods
```{r echo=TRUE}
bw.CvL(childcareSG_ppp.km)


```

```{r echo=TRUE}

bw.scott(childcareSG_ppp.km)

```

```{r echo=TRUE}

bw.ppl(childcareSG_ppp.km)

```

```{r echo=TRUE}
bw.diggle(childcareSG_ppp.km)

```

Compare the output of using bw.diggle and bw.ppl 
``` {r echo=TRUE}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```

### Working with different kernel methods
```{r echo=TRUE}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```


### Fixed and adaptive KDE
```{r echo=TRUE}
kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG_600)

```

Computing KDE by using adaptive bandwidth
```{r echo=TRUE}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)

```

Compare the fixed and adaptive kernel density estimation outputs
```{r echo=TRUE}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")

```

### Converting KDE output into grid object
```{r echo=TRUE}
gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame.im(kde_childcareSG.bw)
spplot(gridded_kde_childcareSG_bw)

```

Converting gridded output into raster
```{r echo=TRUE}
kde_childcareSG_bw_raster <- raster(gridded_kde_childcareSG_bw)

kde_childcareSG_bw_raster
```

Assigning Projection systems
```{r echo=TRUE}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster

```

### Visualising the output in tmap
```{r echo=TRUE}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)

```

### Comparing Spatial Point Patterns using KDE
```{r echo=TRUE}
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
tm = mpsz[mpsz@data$PLN_AREA_N == "TAMPINES",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]

par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")

```

Converting the spatial point dataframe into generic SP format
```{r echo=TRUE}
pg_sp = as(pg, "SpatialPolygons")
tm_sp = as(tm, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
```

Creating owin object
```{r echo=TRUE}
pg_owin = as(pg_sp, "owin")
tm_owin = as(tm_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")

```

Combining childcare points and study area
```{r echo=TRUE}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]

```

Rescale the unit of measurement from metre to kilometer
```{r echo=TRUE}
childcare_pg_ppp.km = rescale(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale(childcare_jw_ppp, 1000, "km")


```

Plot the four areas of study
```{r echo=TRUE}
par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")


```


Computing KDE
```{r echo=TRUE}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")

```


Computing Fixed Bandwidth KDE
```{r echo=TRUE}
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")

```


### Nearest Neighbours Analysis
```{r echo=TRUE}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)

```

Clark and Evans Test: Choa Chu Kang planning area
```{r echo=TRUE}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)

```

Clark and Evans Test: Tampines planning area
```{r echo=TRUE}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)

```



## Second-order Spatial Point Patterns Analysis


Choa Chua Kang Planning Area
Computing G-FUNCTION Estimation
```{r echo=TRUE}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))


```

Performing complete spatial randomness test
```{r echo=TRUE}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)

```

Plot the curve
```{r echo=TRUE}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)

```

Tampines Planning area
Computing G-FUNCTION Estimation
```{r echo=TRUE}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)


```

Performing complete spatial randomness test
```{r echo=TRUE}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)

```

Plot the curve
```{r echo=TRUE}
plot(G_tm.csr)

```


## Analysing Spatial Point Process Using F function

Choa Chu Kang Planning Area
COmputing F-function estimation
```{r echo=TRUE}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

Performing Complete Spatial Randomness Test
```{r echo=TRUE}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)

```
Plot the Curve
```{r echo=TRUE}
plot(F_CK.csr)

```

Tampines Planning Area
COmputing F-function estimation
Monte Carlo test with F-function
```{r echo=TRUE}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

Performing Complete Spatial Randomness Test
```{r echo=TRUE}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)

```
Plot the Curve
```{r echo=TRUE}
plot(F_tm.csr)

```


## Analysing Spatial Point Process Using k-function

Choa Chu Kang Planning area
COmputing k-function estimate
```{r echo=TRUE}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")

```

Performing Complete Spatial Randomness Test
```{r echo=TRUE}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)

```

Plot the Curve
```{r echo=TRUE}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")

```


Tampines Planning Area
COmputing k-function estimate
```{r echo=TRUE}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))

```

Performing Complete Spatial Randomness Test
```{r echo=TRUE}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)

```

Plot the Curve
```{r echo=TRUE}
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))

```




## Analysing Spatial Point Process Using L Function
Choa Chu Kang

Computing L Function Estimation
```{r echo=TRUE}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")

```

Performing COmplete Spatial Randomness Test
```{r echo=TRUE}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

Plot the Curve
```{r echo=TRUE}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")

```


Tampines Planning Area
Computing L Function Estimation
```{r echo=TRUE}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))

```

Performing COmplete Spatial Randomness Test
```{r echo=TRUE}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

Plot the Curve
```{r echo=TRUE}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))

```






























